Unsupervised Detection and Tracking of Multiple Objects with Dependent Dirichlet Process Mixtures

### 2012-01-11 ltx/paper.tex
some more edits

We define the state of the sampler to consist of the assignment variables for all observations, $c_{1:N_{1},1}, \ldots, c_{1:N_{T}, T}$, the deletion variables for all observations, $d_{1:N_{1},1}, \ldots, d_{1:N_{T}, T}$, the sequence of cluster parameters for each active cluster, $\theta_{k,1}, \ldots, \theta_{k, T}$ and the $M$ associated auxiliary variables for each cluster $\theta_{k,t}$ at time $t$, $z_{k,t,1}, \ldots, z_{k, t, M}$. Formally, we can write
\begin{equation}
\text{State} = \{ \bold{c}_{1:T}, \bold{d}_{1:T}, \Theta_{1:T}, \bold{m}_{1:T} \}
\end{equation}
where $\bold{c}_{t} = c_{1:N_{t},t}$, $\bold{d}_{t} = d_{1:N_{t},t}, \Theta_{t} = \theta_{1:K_{t}, t}$, and $m_{1:K_{t},t}$.
The sampler performs an MCMC method known as Gibbs sampling; it operates by iteratively sampling each member of the state given all other members. In particular, at each time $t$ it moves sequentially through each of the $N_{t}$ observations, sampling the assignment $c_{i,t}$ and the deletion variable $d_{i,t}$. After the final observation at a given time is sampled, the cluster parameters for all active clusters at $t$ are sampled, and a M-H step is employed to sample the $M$ auxiliary variables for each active cluster. The following sections detail how each sample in the model is generated.

@@

A value proportional to the posterior probability can be computed for each possible value that $c_{i,t}$ may take on; these values allow us to construct a discrete probability distribution that may be drawn from to get a sample from the posterior distribution over assignments. The possible values that the assignment may take on are $c_{i,t} = k \in \{ 1 , \ldots ,  K_{t}^{'}, k^{*}\}$, where $K_{t}^{'}$ is the number clusters with a non-zero size at any point between time $t$ and the assignment's associated deletion time $d_{i,t}$, and $k^{*}$ is a potential new cluster. The probability that $c_{i,t}$ is assigned to a cluster $k$ can be computed for each $k \in \{ 1 , \ldots ,  K_{t}^{'}, k^{*} \}$, given all other members of the state, by
\begin{equation}
\begin{split}
p(c_{i,t} = k | \text{State}) & \propto
\prod_{i'=i}^{N_{t}}  \text{CRP}(c_{i',t} | \bold{m}_{t-1}, \alpha) \\
& \times \prod_{t' = t+1}^{d_{i,t}}  \prod_{i'=1}^{N_{t'}}   \text{CRP}(c_{i',t'} | \bold{m}_{t'-1}, \alpha) \\
 & \times
\begin{cases}
	P(\bold{x}_{i,t} | \theta_{k,t})                                        \hspace{12pt} \text{if} \hspace{4pt} k \in \{ 1, \ldots, K_{t}^{'} \} \\
	\int P(\bold{x}_{i,t} | \theta) \mathbb{G}_{0}(\theta) d\theta    \hspace{12pt} \text{if}  \hspace{4pt}  k = k^{*}
\end{cases}
\end{split}
\end{equation}
% where $t$ is the time of observation $i$, and the distribution under the product $P(c_{i'} | m_{1:K_{t}^{'}, i'-1}, c_{1:i'-1})$ is given by 
%\begin{equation}
%P(c_{i} = k | m_{1:K_{t}^{'}, i-1}, c_{1:i-1}, \alpha) \propto 
%\begin{cases}
%m_{k, i-1} \hspace{12pt} \text{if} \hspace{4pt} k \in \{ 1, \ldots, K_{t}^{'} \} \\
%\alpha \hspace{12pt} \text{if}  \hspace{4pt}  k = k^{*}
%\end{cases}
%\end{equation}
where the distribution CRP is given by \eqref{crp_rep}, and the sizes of the clusters $\bold{m}_{t'-1}$ are calculated given the fact that $c_{i,t} = k$. Note that the above integral can be determined analytically to give
\begin{equation}
\label{marginal_obs_prob}
\int P(\bold{x}_{i} | \theta) \mathbb{G}_{0}(\theta)d\theta = \text{Student}_{\nu_{n} - 1}(\mu_{n}, \Lambda_{n}/(\kappa_{n}(\nu_{n}-1)))
\end{equation}
where ``Student'' denotes the Student's t distribution, where we follow the three-value parameterization (location parameter, scale parameter, and degrees of freedom) given in \cite{gelman2004bayesian}.

%%%%%
% Need to include ratio of Dirichlet also in the above posterior predictive distro of a single observation
%%%%%

@@

\begin{algorithm}[h!]
\label{alg:MH}
\caption{Metropolis Hastings Algorithm}
\begin{algorithmic}[1]
\STATE $f(x)$ is the stationary PDF from which samples are desired \hfill $\triangleright$ eq. \eqref{stationary_pdf}
\STATE $q( \cdot | x_{i-1})$ is the proposal pdf \hfill $\triangleright$ eq. \eqref{proposal_distro}
\STATE $x_{0} \leftarrow$ initialize
\STATE B $\leftarrow$ \# of burn in samples
\FOR{$i = 1 : \text{B} + N$}
\STATE Sample $x^{*} \sim q( \cdot | x_{i-1})$
\STATE Sample $\alpha \sim \text{Uniform}(0,1)$
\STATE $r_{accept} \leftarrow \frac{f(x^{*})q(x_{i-1} | x^{*})} {f(x_{i-1})q(x^{*} | x_{i-1})} $ \hfill $\triangleright$  eq. \eqref{accept_ratio}
\IF{$\alpha < \text{min} \{ 1, r_{accept} \} $}
\STATE $x_{i} \leftarrow x^{*}$
\ELSE
\STATE $x_{i} \leftarrow x_{i-1}$
\ENDIF
\ENDFOR
\STATE Discard samples $x_{1}, \ldots, x_{\text{B}}$
\STATE \textbf{Output}: Remaining samples $x_{1}, \ldots, x_{N}$
\end{algorithmic}
%\caption{Say thing at end}
\end{algorithm}

The M-H algorithm is used to sample auxiliary variables for each alive cluster. In short, we sample a new value for all $M$ auxiliary variables, denoted $z_{k,t,m}^{*}$, according to the following proposal distribution
\begin{equation}
\label{proposal_distro}
z_{k,t,m}^{*}  \sim  P(z_{k,t,m} | \theta_{k,t}) = \mathcal{N}(z_{k,t,m}^{s} | \boldsymbol{\mu}_{k,t}, \Sigma_{k,t}) \mathcal{M}n(z_{k,t,m}^{c} | \bold{p}_{k,t})
\end{equation}
compute the standard M-H acceptance ratio, which in this case (due to the proposal distribution choice) simplifies to 
\begin{equation}
\label{accept_ratio}
r_{accept} = \frac{P(\theta_{k,t-1} | z_{k,t,m}^{*})}{P(\theta_{k,t-1} | z_{k,t,m})}
\end{equation}
and keep this sample with probability $\text{min}(1, r_{accept})$.

@@

\subsubsection{Sampling Deletion Variables $d_{i}$}
Computing the posterior distribution (or a value proportional to this distribution) for each possible value that a given deletion variable may take on (in a manner similar to how samples from the posterior distribution of the assignment variables $c_{i}$ are gained in Section~\ref{sec:sample_assignments}) may be computationally expensive due to the large number of potential deletion times. To remedy this, the M-H algorithm may again be used to sample from the posterior, where the prior distribution for the deletion times can be used to propose new samples $d_{i}^{*}$. As the `lifetime' of each assignment can be shown to be geometrically distributed, we can propose a new deletion time sample with
\begin{align}
\begin{split}
l_{i}  &\sim  \text{Geo}(\rho)  \\
d_{i}^{*}  &= l_{i} + t + 1
\end{split}
\end{align}
and accept or reject this sample using the standard M-H acceptance ratio.

@@

% mcmc: initializing the sampler
% ------------------------------

\subsubsection{Initializing the Sampler}
Before the samples can be iteratively drawn from each element of the state given all other elements, the state must be initialized. While the state of the sampler can be initialized at any value, beginning at a value closer to the maximum joint probability (over the state) allows for quicker inference.

@@

This section details a Sequential Monte Carlo (SMC) sampler---also known as a particle filter---used to perform inference, which provides another method to carry out unsupervised detection and tracking. The SMC inference algorithm is given in Algorithm~\ref{alg:SMC}. In this algorithm, at each time step $t \in \{ 1, \ldots, T \}$, a number of samples refered to as ``particles'' are generated; each particle consists of a sample from the posterior distribution over the assignment for each observation, $c_{1,t}, \ldots, c_{N_{t}, t}$, parameters for each cluster $\theta_{1,t}, \ldots, \theta_{K_{t}, t}$, and size after deletion for each cluster $m_{1,t}, \ldots, m_{K_{t},t}$. A set of particles is sampled at each time step from relevant proposal distributions (described in Sections~\ref{sec:smc_proposal_1}, \ref{sec:smc_proposal_2}, and \ref{sec:smc_proposal_3}), a weight is computed for each particle, and a subset of the particles are resampled and kept via a resampling process. Within each time step, Gibbs sampling is used to generate the samples associated with each particle.

The sequence of target distributions for the SMC algorithm may be written as
\begin{equation}
\begin{split}
%\pi_{t}(c_{1:N_{1}, 1}, \ldots, c_{1:N_{t}, t}, & \theta_{1:K_{1}, 1}, \ldots, \theta_{1:K_{t}, t}, m_{1:K_{1}, 1}, \ldots, m_{1:K_{t}, t}) = \\
%\pi_{t-1}(c_{1:N_{1}, 1}, \ldots, c_{1:N_{t-1}, t-1}, & \theta_{1:K_{1}, 1}, \ldots, \theta_{1:K_{t-1}, t-1}, m_{1:K_{1}, 1}, \ldots, m_{1:K_{t-1}, t-1}) \\
\pi_{t}(\bold{c}_{1:t}, \Theta_{1:t}, \bold{m}_{1:t}) & =  \pi_{t-1}(\bold{c}_{1:t-1}, \Theta_{1:t-1}, \bold{m}_{1:t-1})\\
& \times \prod_{i=1}^{N_{t}} P(c_{i,t} | \bold{m}_{t}, \Theta_{t}, \bold{c}_{1:t}, \bold{x}_{1:N_{t}}) \\
& \times \prod_{k=1}^{K} 
\begin{cases}
\begin{split}
P(\theta_{k,t} | \theta_{k,t-1}) \hspace{2mm} & \text{if} \hspace{2mm} k \leq K_{t-1} \\
\mathbb{G}_{0} \hspace{18mm} & \text{if} \hspace{2mm} k > K_{t-1}
\end{split}
\end{cases} \\
& \times \prod_{k=1}^{K} \text{DEL}(m_{k,t} | m_{k,t-1}, c_{1:N_{t-1}, t-1}, \rho)
\end{split}
\end{equation}
where $\bold{c}_{t} = c_{1,t}, \ldots, c_{N_{t}, t}$, $\Theta_{t} = \theta_{1,t}, \ldots, \theta_{K_{t}, t}$, and $\bold{m}_{t} = m_{1,t}, \ldots, m_{K_{t}, t}$ A few distributions referenced in the SMC algorithm are described in detail in the following sections.

@@

\begin{algorithm}[!]
\caption{Sequential Monte Carlo Inference for the GPUDDPM}
\label{alg:SMC}
\begin{algorithmic}[1]
\FOR{$l = 1 : L$}
\STATE $w_{0}^{(l)} \leftarrow 1/L$ \hfill $\triangleright$ initialize weights
\ENDFOR
\STATE $K_{0}^{(l)} \leftarrow 0$ \hfill $\triangleright$ initialize \# of clusters
\FOR{$t = 1 : T  \text{  (\# of frames})$}
\FOR{$l = 1 : L  \text{  (\# of particles})$}
\STATE $K_{t}^{(l)}  \leftarrow K_{t-1}^{(l)}$
\STATE $m_{1:K_{t}^{(l)}, t}^{(l)} \leftarrow m_{1:K_{t-1}^{(l)}, t-1}^{(l)}$
\FOR{$s = 1 : S \text{  (\# of Gibbs samples})$}
\FOR{$i = 1 : N_{t} \text{  (\# of observations at frame $t$})$}
\IF{$s=1$}
\STATE Sample $c_{i,t}^{(l)} \sim P \left( c_{i,t}^{(l)} | m_{1:K_{t-1}^{(l)}, t-1}^{(l)}, \theta_{1:K_{t-1}^{(l)}, t-1}^{(l)}, \alpha \right)$  \hfill $\triangleright$ eq.~\eqref{smc_assig_eqn}
\STATE $m_{c_{i,t}^{(l)},t}^{(l)} \leftarrow m_{c_{i,t}^{(l)},t}^{(l)} + 1$
\ELSE
\STATE $m_{c_{i,t}^{(l)},t}^{(l)} \leftarrow m_{c_{i,t}^{(l)},t}^{(l)} - 1$
\STATE Sample $c_{i,t}^{(l)} \sim P \left( c_{i,t}^{(l)} | m_{1:K_{t}^{(l)}, t}^{(l)}, \theta_{1:K_{t}^{(l)}, t}^{(l)}, \alpha \right)$  \hfill $\triangleright$ eq.~\eqref{smc_assig_eqn}
\STATE $m_{c_{i,t}^{(l)},t}^{(l)} \leftarrow m_{c_{i,t}^{(l)},t}^{(l)} + 1$
\ENDIF
\IF{$c_{i,t}^{(l)} = K_{t}^{(l)} + 1  \text{  (a new cluster)}$}
\STATE Sample $\theta_{c_{i,t}^{(l)}, t}^{(l)} \sim q_{1}(\bold{x}_{i,t})$ \hfill $\triangleright$ eq.~\eqref{smc_q1_eqn}
\STATE $K_{t}^{(l)} \leftarrow K_{t}^{(l)} + 1$
\STATE $m_{K_{t}^{(l)},t}^{(l)} \leftarrow 1$
\ENDIF
\ENDFOR
\FOR{$k = 1 : K_{t}^{(l)}  \text{  (\# of clusters at frame $t$})$}
\IF{$k > K_{t-1}^{(l)}$}
\STATE Sample $\theta_{k,t}^{(l)} \sim q_{1}(\{ \bold{x}_{1:N_{t}, t} = k \})$  \hfill $\triangleright$ eq.~\eqref{smc_q1_eqn}
\ELSIF{$k \leq K_{t-1}^{(l)}$ and $ \#\{ \bold{x}_{1:N_{t}, t} = k \} > 0$}
\STATE Sample $\theta_{k,t}^{(l)} \sim q_{2}(\theta_{k,t-1}^{(l)}, \{ \bold{x}_{1:N_{t}, t} = k \})$  \hfill $\triangleright$ eq.~\eqref{smc_q2_eqn}
\ELSIF{$m_{k,t}^{(l)} > 0$}
\STATE Sample $\theta_{k,t}^{(l)} \sim P(\theta_{k,t}^{(l)} | \theta_{k,t-1}^{(l)})$  \hfill $\triangleright$ eqs.~\eqref{transker1} $\&$ \eqref{transker2}
\ENDIF
\IF{$s=S$}
\STATE Sample $m_{k,t+1}^{(l)} \sim \text{DEL}(m_{k,t}^{(l)}, c_{1:N_{t}, t}^{(l)}, \rho)$  \hfill $\triangleright$ eq.~\eqref{del_step}
\ENDIF
\ENDFOR
\ENDFOR
\STATE $\tilde{w}_{t}^{(l)} \leftarrow w_{t-1}^{(l)} \times \frac{ P(\bold{x}_{1:N_{t},t}, c_{1:N_{t}, t}^{(l)} | \theta_{1:K_{t}^{(l)}}, m_{1:K_{t},t}^{(l)})}	{P(c_{1:N_{t}, t}^{(l)} | m_{t}^{(l)}, \theta_{t-1}^{(l)}, \bold{x}_{1:N_{t},t} ) } \times r $  \hfill $\triangleright$ eq.~\eqref{smc_weight_eqn}
\ENDFOR
\FOR {l = 1 : L}
\STATE $w_{t}^{(l)} \leftarrow \frac{\tilde{w}_{t}^{(l)}}{\sum_{l=1}^{L} \tilde{w}_{t}^{(l)} }$   \hfill $\triangleright$ normalize weights
\ENDFOR
\STATE Resample particles $1, \ldots, L$ and weights $w^{(1)}, \ldots, w^{(L)}$   \hfill $\triangleright$ Section~\ref{sec:resample}
\ENDFOR
\end{algorithmic}
\end{algorithm}

@@

The probability of assignments given current cluster sizes, cluster parameters, and the Dirichlet Process granularity parameter $\alpha$ can be written as
\begin{equation}
\label{smc_assig_eqn}
\begin{split}
P \left( c_{i,t} | m_{1:K_{t}, t}, \theta_{1:K_{t}, t}, \alpha \right) &\propto
\text{CRP}(m_{1:K_{t}, t}, \alpha) \\ &\times
\begin{cases}
P(\bold{x}_{i,t} | \theta_{c_{i,t},t}) \hspace{2mm} \text{if} \hspace{2mm} k \leq K_{t-1} \\
\int P(\bold{x}_{i,t} | \theta) \mathbb{G}_{0}(\theta)d\theta \hspace{2mm} k > K_{t-1}
\end{cases}
\end{split}
\end{equation}
The integral can $\int P(\bold{x}_{i,t} | \theta) \mathbb{G}_{0}(\theta)d\theta$ can be determined analytically, and is given in \eqref{marginal_obs_prob}.

@@

At each time step $t \in \{ 1, \ldots, T \}$, after all of the $L$ particles have been sampled and their associated weights computed, a resampling step is carried out. In this step, $L$ new particles are sampled from currnt set of $L$ particles.

>>>>

We define the state of the sampler to consist of the assignment variables for all observations, $c_{1:N_{1},1}, \ldots, c_{1:N_{T}, T}$, the deletion variables for all observations, $d_{1:N_{1},1}, \ldots, d_{1:N_{T}, T}$, the sequence of cluster parameters for each active cluster, $\theta_{k,1}, \ldots, \theta_{k, T}$ and the $M$ associated auxiliary variables for each cluster $\theta_{k,t}$ at time $t$, $z_{k,t,1}, \ldots, z_{k, t, M}$. Formally, we can write
\begin{equation}
\text{State} = \{ \bold{c}_{1:T}, \bold{d}_{1:T}, \Theta_{1:T}, \bold{m}_{1:T} \}
\end{equation}
where $\bold{c}_{t} = c_{1:N_{t},t}$, $\bold{d}_{t} = d_{1:N_{t},t}, \Theta_{t} = \theta_{1:K_{t}, t}$, and $m_{1:K_{t},t}$.
The sampler performs an MCMC method known as Gibbs sampling; it operates by iteratively sampling each member of the state given all other members. In particular, at each time $t$ it moves sequentially through each of the $N_{t}$ observations, sampling the assignment $c_{i,t}$ and the deletion variable $d_{i,t}$. After the final observation at a given time is sampled, the cluster parameters for all active clusters at $t$ are sampled, and a M-H step is employed to sample the $M$ auxiliary variables for each active cluster. Before the samples can be iteratively drawn from each element of the state given all other elements, the state must be initialized. While the state of the sampler can be initialized at any value, beginning at a value closer to the maximum joint probability (over the state) allows for quicker inference. The following sections detail how each sample in the model is drawn.

@@

A value proportional to the posterior probability can be computed for each possible value that $c_{i,t}$ may take on; these values allow us to construct a discrete probability distribution from which we can draw samples from the posterior distribution over assignments. The possible values that the assignment may take on are $c_{i,t} = k \in \{ 1 , \ldots ,  K_{t}^{'}, k^{*}\}$, where $K_{t}^{'}$ is the number clusters with a non-zero size at any point between time $t$ and the assignment's associated deletion time $d_{i,t}$, and $k^{*}$ is a potential new cluster. The probability that $c_{i,t}$ is assigned to a cluster $k$ can be computed for each $k \in \{ 1 , \ldots ,  K_{t}^{'}, k^{*} \}$, given all other members of the state, by
\begin{equation}
\begin{split}
p(c_{i,t} = k | \text{State}) & \propto
\prod_{i'=i}^{N_{t}}  \text{CRP}(c_{i',t} | \bold{m}_{t-1}, \alpha) \\
& \times \prod_{t' = t+1}^{d_{i,t}}  \prod_{i'=1}^{N_{t'}}   \text{CRP}(c_{i',t'} | \bold{m}_{t'-1}, \alpha) \\
 & \times
\begin{cases}
	P(\bold{x}_{i,t} | \theta_{k,t})                                        \hspace{12pt} \text{if} \hspace{4pt} k \in \{ 1, \ldots, K_{t}^{'} \} \\
	\int P(\bold{x}_{i,t} | \theta) \mathbb{G}_{0}(\theta) d\theta    \hspace{12pt} \text{if}  \hspace{4pt}  k = k^{*}
\end{cases}
\end{split}
\end{equation}
% where $t$ is the time of observation $i$, and the distribution under the product $P(c_{i'} | m_{1:K_{t}^{'}, i'-1}, c_{1:i'-1})$ is given by 
%\begin{equation}
%P(c_{i} = k | m_{1:K_{t}^{'}, i-1}, c_{1:i-1}, \alpha) \propto 
%\begin{cases}
%m_{k, i-1} \hspace{12pt} \text{if} \hspace{4pt} k \in \{ 1, \ldots, K_{t}^{'} \} \\
%\alpha \hspace{12pt} \text{if}  \hspace{4pt}  k = k^{*}
%\end{cases}
%\end{equation}
where the distribution CRP is given by \eqref{crp_rep}, and the sizes of the clusters $\bold{m}_{t'-1}$ are calculated given the fact that $c_{i,t} = k$. Note that the above integral can be determined analytically to give
\begin{equation}
\label{marginal_obs_prob}
\int P(\bold{x}_{i} | \theta) \mathbb{G}_{0}(\theta)d\theta = \text{Student}_{\nu_{n} - 1}(\mu_{n}, \Lambda_{n}/(\kappa_{n}(\nu_{n}-1)))
\end{equation}
where ``Student'' denotes the Student's t distribution, where we follow the three-value parameterization (location parameter, scale parameter, and degrees of freedom) given in \cite{gelman2004bayesian}.

%%%%%
% Need to include ratio of Dirichlet also in the above posterior predictive distro of a single observation, and discuss how to compute any parameters in here that aren't yet specified
%%%%%

@@

\begin{algorithm}[h!]
\label{alg:MH}
\caption{Metropolis Hastings Algorithm}
\begin{algorithmic}[1]
\STATE $f(x)$ is the stationary PDF from which $N$ samples are desired \hfill $\triangleright$ eq. \eqref{stationary_pdf}
\STATE $q( \cdot | x_{i-1})$ is the proposal pdf \hfill $\triangleright$ eq. \eqref{proposal_distro}
\STATE B is the number of burn in samples
\STATE $x_{0} \leftarrow$ initialize
%\STATE B $\leftarrow$ \# of burn in samples
\FOR{$i = 1 : \text{B} + N$}
\STATE Sample $x^{*} \sim q( \cdot | x_{i-1})$
\STATE Sample $\alpha \sim \text{Uniform}(0,1)$
\STATE $r_{accept} \leftarrow \frac{f(x^{*})q(x_{i-1} | x^{*})} {f(x_{i-1})q(x^{*} | x_{i-1})} $ \hfill $\triangleright$  eq. \eqref{accept_ratio}
\IF{$\alpha < \text{min} \{ 1, r_{accept} \} $}
\STATE $x_{i} \leftarrow x^{*}$
\ELSE
\STATE $x_{i} \leftarrow x_{i-1}$
\ENDIF
\ENDFOR
\STATE Discard samples $x_{1}, \ldots, x_{\text{B}}$
\STATE \textbf{Output}: Remaining samples $x_{1}, \ldots, x_{N}$
\end{algorithmic}
%\caption{Say thing at end}
\end{algorithm}

The M-H algorithm is used to sample auxiliary variables for each alive (positive size) cluster. In short, we sample a new value for all $M$ auxiliary variables, denoted $z_{k,t,m}^{*}$, according to the following proposal distribution
\begin{equation}
\label{proposal_distro}
z_{k,t,m}^{*}  \sim  P(z_{k,t,m} | \theta_{k,t}) = \mathcal{N}(z_{k,t,m}^{s} | \boldsymbol{\mu}_{k,t}, \Sigma_{k,t}) \mathcal{M}n(z_{k,t,m}^{c} | \bold{p}_{k,t})
\end{equation}
compute the standard M-H acceptance ratio, which in this case (due to the proposal distribution choice) simplifies to 
\begin{equation}
\label{accept_ratio}
r_{accept} = \frac{P(\theta_{k,t-1} | z_{k,t,m}^{*})}{P(\theta_{k,t-1} | z_{k,t,m})}
\end{equation}
and choose these auxiliary variables as the next sample in the constructed sequence if $\alpha < \text{min}(1, r_{accept})$, where $\alpha$ is a sample from a $\text{Uniform}(0,1)$ distribution.

@@

\subsubsection{Sampling Deletion Variables $d_{i}$}
Computing the posterior distribution (or a value proportional to this distribution) for each possible value that a given deletion variable may take on, may be accomplished in a manner similar to how samples from the posterior distribution of the assignment variables $c_{i}$ are gained in Section~\ref{sec:sample_assignments}; however, this may be computationally expensive due to the large number of potential deletion times. To remedy this, the M-H algorithm may again be used to sample from the posterior, where the prior distribution for the deletion times can be used to propose new samples $d_{i}^{*}$. As the ``lifetime'' of each assignment can be shown to be geometrically distributed, we can propose a new deletion time sample with
\begin{align}
\begin{split}
l_{i}  &\sim  \text{Geo}(\rho)  \\
d_{i}^{*}  &= l_{i} + t + 1
\end{split}
\end{align}
and accept or reject this sample using the standard M-H acceptance ratio.

@@

@@

This section details a Sequential Monte Carlo (SMC) sampler---also known as a particle filter---used to perform inference, which provides another method to carry out unsupervised detection and tracking. The SMC inference algorithm is given in Algorithm~\ref{alg:SMC}. In this algorithm, at each time step $t \in \{ 1, \ldots, T \}$, a number of samples refered to as ``particles'' are generated; each particle consists of a sample from the posterior distribution over the assignment for each observation, $c_{1,t}, \ldots, c_{N_{t}, t}$, parameters for each cluster $\theta_{1,t}, \ldots, \theta_{K_{t}, t}$, and size after deletion for each cluster $m_{1,t}, \ldots, m_{K_{t},t}$. A set of particles is sampled at each time step from relevant proposal distributions (described in Sections~\ref{sec:smc_proposal_1}, \ref{sec:smc_proposal_2}, and \ref{sec:smc_proposal_3}), a weight is computed for each particle, and a new set of particles are sampled from the set of weighted particles via a resampling process. Within each time step, Gibbs sampling is used to generate the samples associated with each particle.

The sequence of target distributions for the SMC algorithm may be written as
\begin{equation}
\begin{split}
%\pi_{t}(c_{1:N_{1}, 1}, \ldots, c_{1:N_{t}, t}, & \theta_{1:K_{1}, 1}, \ldots, \theta_{1:K_{t}, t}, m_{1:K_{1}, 1}, \ldots, m_{1:K_{t}, t}) = \\
%\pi_{t-1}(c_{1:N_{1}, 1}, \ldots, c_{1:N_{t-1}, t-1}, & \theta_{1:K_{1}, 1}, \ldots, \theta_{1:K_{t-1}, t-1}, m_{1:K_{1}, 1}, \ldots, m_{1:K_{t-1}, t-1}) \\
\pi_{t}(\bold{c}_{1:t}, \Theta_{1:t}, \bold{m}_{1:t}) & =  \pi_{t-1}(\bold{c}_{1:t-1}, \Theta_{1:t-1}, \bold{m}_{1:t-1})\\
& \times \prod_{i=1}^{N_{t}} P(c_{i,t} | \bold{m}_{t}, \Theta_{t}, \bold{c}_{1:t}, \bold{x}_{1:N_{t}}) \\
& \times \prod_{k=1}^{K} 
\begin{cases}
\begin{split}
P(\theta_{k,t} | \theta_{k,t-1}) \hspace{2mm} & \text{if} \hspace{2mm} k \leq K_{t-1} \\
\mathbb{G}_{0} \hspace{18mm} & \text{if} \hspace{2mm} k > K_{t-1}
\end{split}
\end{cases} \\
& \times \prod_{k=1}^{K} \text{DEL}(m_{k,t} | m_{k,t-1}, c_{1:N_{t-1}, t-1}, \rho)
\end{split}
\end{equation}
where $\bold{c}_{t} = c_{1,t}, \ldots, c_{N_{t}, t}$, $\Theta_{t} = \theta_{1,t}, \ldots, \theta_{K_{t}, t}$, and $\bold{m}_{t} = m_{1,t}, \ldots, m_{K_{t}, t}$ The distributions used in the SMC algorithm are described in detail in the following sections.

@@

\begin{algorithm}[!]
\caption{Sequential Monte Carlo Inference for the GPUDDPM}
\label{alg:SMC}
\begin{algorithmic}[1]
\FOR{$l = 1 : L$}
\STATE $w_{0}^{(l)} \leftarrow 1/L$ \hfill $\triangleright$ initialize weights
\ENDFOR
\STATE $K_{0}^{(l)} \leftarrow 0$ \hfill $\triangleright$ initialize \# of clusters
\FOR{$t = 1 : T  \text{  (\# of frames})$}
\FOR{$l = 1 : L  \text{  (\# of particles})$}
\STATE $K_{t}^{(l)}  \leftarrow K_{t-1}^{(l)}$
\STATE $m_{1:K_{t}^{(l)}, t}^{(l)} \leftarrow m_{1:K_{t-1}^{(l)}, t-1}^{(l)}$
\FOR{$s = 1 : S \text{  (\# of Gibbs samples})$}
\FOR{$i = 1 : N_{t} \text{  (\# of observations at frame $t$})$}
\IF{$s=1$}
\STATE Sample $c_{i,t}^{(l)} \sim P \left( c_{i,t}^{(l)} | m_{1:K_{t-1}^{(l)}, t-1}^{(l)}, \theta_{1:K_{t-1}^{(l)}, t-1}^{(l)}, \alpha \right)$  \hfill $\triangleright$ eq.~\eqref{smc_assig_eqn}
\STATE $m_{c_{i,t}^{(l)},t}^{(l)} \leftarrow m_{c_{i,t}^{(l)},t}^{(l)} + 1$
\ELSE
\STATE $m_{c_{i,t}^{(l)},t}^{(l)} \leftarrow m_{c_{i,t}^{(l)},t}^{(l)} - 1$
\STATE Sample $c_{i,t}^{(l)} \sim P \left( c_{i,t}^{(l)} | m_{1:K_{t}^{(l)}, t}^{(l)}, \theta_{1:K_{t}^{(l)}, t}^{(l)}, \alpha \right)$  \hfill $\triangleright$ eq.~\eqref{smc_assig_eqn}
\STATE $m_{c_{i,t}^{(l)},t}^{(l)} \leftarrow m_{c_{i,t}^{(l)},t}^{(l)} + 1$
\ENDIF
\IF{$c_{i,t}^{(l)} = K_{t}^{(l)} + 1  \text{  (a new cluster)}$}
\STATE Sample $\theta_{c_{i,t}^{(l)}, t}^{(l)} \sim q_{1}(\bold{x}_{i,t})$ \hfill $\triangleright$ eq.~\eqref{smc_q1_eqn}
\STATE $K_{t}^{(l)} \leftarrow K_{t}^{(l)} + 1$
\STATE $m_{K_{t}^{(l)},t}^{(l)} \leftarrow 1$
\ENDIF
\ENDFOR
\FOR{$k = 1 : K_{t}^{(l)}  \text{  (\# of clusters at frame $t$})$}
\IF{$k > K_{t-1}^{(l)}$}
\STATE Sample $\theta_{k,t}^{(l)} \sim q_{1}(\{ \bold{x}_{1:N_{t}, t} = k \})$  \hfill $\triangleright$ eq.~\eqref{smc_q1_eqn}
\ELSIF{$k \leq K_{t-1}^{(l)}$ and $ \#\{ \bold{x}_{1:N_{t}, t} = k \} > 0$}
\STATE Sample $\theta_{k,t}^{(l)} \sim q_{2}(\theta_{k,t-1}^{(l)}, \{ \bold{x}_{1:N_{t}, t} = k \})$  \hfill $\triangleright$ eq.~\eqref{smc_q2_eqn}
\ELSIF{$m_{k,t}^{(l)} > 0$}
\STATE Sample $\theta_{k,t}^{(l)} \sim P(\theta_{k,t}^{(l)} | \theta_{k,t-1}^{(l)})$  \hfill $\triangleright$ eqs.~\eqref{transker1} $\&$ \eqref{transker2}
\ENDIF
\IF{$s=S$}
\STATE Sample $m_{k,t+1}^{(l)} \sim \text{DEL}(m_{k,t}^{(l)}, c_{1:N_{t}, t}^{(l)}, \rho)$  \hfill $\triangleright$ eq.~\eqref{del_step}
\ENDIF
\ENDFOR
\ENDFOR
\STATE $\tilde{w}_{t}^{(l)} \leftarrow w_{t-1}^{(l)} \times \frac{ P(\bold{x}_{1:N_{t},t}, c_{1:N_{t}, t}^{(l)} | \theta_{1:K_{t}^{(l)}}, m_{1:K_{t},t}^{(l)})}	{P(c_{1:N_{t}, t}^{(l)} | m_{t}^{(l)}, \theta_{t-1}^{(l)}, \bold{x}_{1:N_{t},t} ) } \times r $  \hfill $\triangleright$ eq.~\eqref{smc_weight_eqn}
\ENDFOR
\FOR {l = 1 : L}
\STATE $w_{t}^{(l)} \leftarrow \frac{\tilde{w}_{t}^{(l)}}{\sum_{l=1}^{L} \tilde{w}_{t}^{(l)} }$   \hfill $\triangleright$ normalize weights
\ENDFOR
\STATE Resample particles $1, \ldots, L$ and weights $w_{t}^{(1)}, \ldots, w_{t}^{(L)}$   \hfill $\triangleright$ Section~\ref{sec:resample}
\ENDFOR
\end{algorithmic}
\end{algorithm}

@@

The probability of assignments given current cluster sizes, cluster parameters, and the Dirichlet Process granularity parameter $\alpha$ can be written as
\begin{equation}
\label{smc_assig_eqn}
\begin{split}
P \left( c_{i,t} | m_{1:K_{t}, t}, \theta_{1:K_{t}, t}, \alpha \right) &\propto
\text{CRP}(m_{1:K_{t}, t}, \alpha) \\ &\times
\begin{cases}
P(\bold{x}_{i,t} | \theta_{c_{i,t},t}) \hspace{2mm} \text{if} \hspace{2mm} k \leq K_{t-1} \\
\int P(\bold{x}_{i,t} | \theta) \mathbb{G}_{0}(\theta)d\theta \hspace{2mm} k > K_{t-1}
\end{cases}
\end{split}
\end{equation}
where CRP is defined in \eqref{CRP}, and the integral $\int P(\bold{x}_{i,t} | \theta) \mathbb{G}_{0}(\theta)d\theta$ can be determined analytically, and is given in \eqref{marginal_obs_prob}.

@@

At each time step $t \in \{ 1, \ldots, T \}$, after all of the $L$ particles have been sampled and their associated weights computed, a resampling step is carried out. In this step, $L$ new particles are sampled from current set of $L$ particles. We will describe a specific, simple method known as Multinomial Resampling. In this method of resampling, we draw
\begin{equation}
\Upsilon \sim \mathcal{M}n((w_{t}^{(1)}, \ldots, w_{t}^{(L)}))
\end{equation}
where $\mathcal{M}n$ denotes the multinomial distribution, and $\Upsilon$ represents a vector of counts showing the multiplicity of each of the old particles in the new particle set (i.e. $\Upsilon_{i}$ denotes how many times particle $i$ is duplicated in the new particle set). More sophisticated resampling strategies are described for the SMC algorithm in \cite{gasthaus_thesis} and \cite{douc2005comparison}.

%%%%%
% need to include particle resampling strategies
%%%%%
